---
title: "Applied Bayesian Analysis Assignment 2"
author: "Faizan Khalid Mohsin"
date: "February 23, 2019"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
---

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1 

<http://rmarkdown.rstudio.com>.

## Question 1 a:

Done in the attached paper by hand. 

## Question 1 b:

```{r }

# Data

set.seed(25)
nbig<-20000

data = matrix( c( 1.13, 1.75, 2.30, 3.18,
                  1.20, 1.45, 2.15, 3.10,
                  1.00, 1.55 ,2.25, 3.28,
                  0.91, 1.64, 2.40, 3.35,
                  1.05, 1.60, 2.49, 3.12), nrow = 5, byrow = T)

data = as.data.frame(data)
colnames(data) = c("temp40", "temp60", "temp80", "temp100")
attach(data)

Y1<- mean(temp40)
Y2<- mean(temp60)
Y3<- mean(temp80)
Y4<- mean(temp100)

#MCMC Algorithm

mu00<-0
tau00<-1
alp<-1
beta<-1
alp0<-1
beta0<-1
mu1<-rep(0,nbig)
mu2<-rep(0,nbig)
mu3<-rep(0,nbig)
mu4<-rep(0,nbig)
tau<-rep(0,nbig)
mu0post1<-rep(0,nbig) 
tau0post1<-rep(0,nbig) 
mu0post2<-rep(0,nbig)
tau0post2<-rep(0,nbig)
mu0post3<-rep(0,nbig) 
tau0post3<-rep(0,nbig)
mu0post4<-rep(0,nbig) 
tau0post4<-rep(0,nbig) 
mu0<-rep(0,nbig)
tau0<-rep(0,nbig)
mupost00<-rep(0,nbig) 
taupost00<-rep(0,nbig)  
betapost<-rep(0,nbig)
betapost0<-rep(0,nbig)
mu0[1]<-1  
mu1[1]<-0
mu2[1]<-0
mu3[1]<-0
mu4[1]<-0
tau[1]<-1
tau0[1]<-1

for(i in 2:nbig) {
  mu0post1[i]<-(tau0[i-1]*mu0[i-1]+5*tau[i-1]*Y1)/(tau0[i-1]+5*tau[i-1])
  tau0post1[i]<-tau0[i-1]+tau[i-1]*5
  mu1[i]<-rnorm(1,mu0post1[i],sd=1/sqrt(tau0post1[i]))
  mu0post2[i]<-(tau0[i-1]*mu0[i-1]+5*tau[i-1]*Y2)/(tau0[i-1]+5*tau[i-1])
  tau0post2[i]<-tau0[i-1]+tau[i-1]*5
  mu2[i]<-rnorm(1,mu0post2[i],sd=1/sqrt(tau0post2[i]))
  mu0post3[i]<-(tau0[i-1]*mu0[i-1]+5*tau[i-1]*Y3)/(tau0[i-1]+5*tau[i-1])
  tau0post3[i]<-tau0[i-1]+tau[i-1]*5
  mu3[i]<-rnorm(1,mu0post3[i],sd=1/sqrt(tau0post3[i]))
  mu0post4[i]<-(tau0[i-1]*mu0[i-1]+5*tau[i-1]*Y4)/(tau0[i-1]+5*tau[i-1])
  tau0post4[i]<-tau0[i-1]+tau[i-1]*5
  mu4[i]<-rnorm(1,mu0post4[i],sd=1/sqrt(tau0post4[i]))
  mupost00[i]<-(4*tau0[i-1]*((mu1[i-1]+mu2[i-1]+mu3[i-1]+mu4[i-1])/4)+tau00*mu00)/(4*tau0[i-1]+tau00)
  taupost00[i]<-4*tau0[i-1]+ tau00
  mu0[i]<-rnorm(1,mupost00[i],1/sqrt(taupost00[i]))
  alppost0<-alp0+(4/2) 
  betapost0[i]<-beta0+(1/2)*(sum((mu1[i]-mu0[i])^2)+sum((mu2[i]-mu0[i])^2)+
                               sum((mu3[i]-mu0[i])^2)+sum((mu4[i]-mu0[i])^2))
  tau0[i]<-rgamma(1,alppost0,betapost0[i])
  alppost<-alp+(5+5+5+5)/2 
  betapost[i]<- beta+(0.5*(sum((temp40-mu1[i])^2)+sum((temp60-mu2[i])^2)+
                             sum((temp80-mu3[i])^2)+sum((temp100-mu4[i])^2)))
  tau[i]<-rgamma(1,alppost,betapost[i]) 
}

cbind(mu0,mu1,mu2,mu3,mu4,tau,tau0)[1:10,]


```



## Question 1 c:

```{r}
plot(density(mu1[-(1:5)]),type="l", xlab="mu", ylab="Posterior density ",
cex=1.5, xlim=c(0,4), col="red", lty=1, main="Posterior distributions for Each temperature")
lines(density(mu2[-(1:5)]), col="cyan",lty=1)
lines(density(mu3[-(1:5)]),col="black",lty=1)
lines(density(mu4[-(1:5)]),col="blue", lty=1)
legend("topright", c("Temp 40", "Temp 60", "Temp 80", "Temp100"),lty=c(1, 1, 1,1), col=c("red","blue","green","black"),cex=0.5)


```

We now give summary statistics.

```{r}
#some summary statistics 
#for temperature of 40
temp1<-rnorm(nbig,mean=mu1,sd=1/sqrt(tau))
mean(temp1)

sd(temp1)

quantile(temp1, probs=seq(0,1,0.025)) # can use this for quantiles

#for temperature of 60
temp2<-rnorm(nbig,mean=mu2,sd=1/sqrt(tau))
mean(temp2)

sd(temp2)

quantile(temp2, probs=seq(0,1,0.025)) # can use this for quantiles

#for temperature of 80
temp3<-rnorm(nbig,mean=mu3,sd=1/sqrt(tau))
mean(temp3)

sd(temp3)

quantile(temp3, probs=seq(0,1,0.025)) # can use this for quantiles

#for temperature of 100
temp4<-rnorm(nbig,mean=mu4,sd=1/sqrt(tau))
mean(temp4)

quantile(temp4, probs=seq(0,1,0.025)) # can use this for quantiles
sd(temp4)


```


## Question 1 d:

What is the posterior distribution of the difference in number of cells grown at a temperature of 40 versus 80? What is the posterior probability that there will be more cells grown at a temperature of 40 versus 80?

Below we calculate the posterior distribution and probability. 

Posterior probability will be 0, since temp3-temp1 is always less than 0 at all occasions, as indicated by the quantiles and the plot below

```{r}
newmu<-mu1-mu4
mean(temp1-temp3)

sd(temp1-temp3)

quantile(temp1-temp3, probs=seq(0,1,0.025))

```


```{r}
plot(density(temp1-temp3),type="l", xlab="mu difference", ylab="Posterior density ",    cex=1.5, xlim=c(-4,4), col="blue", lty=1, main="Posterior distributions for difference in temp 40 vs 80")
```

# Question 2 

## Question 2 a

```{r}

smokeDat=read.csv("SmokeAgeDeath.csv")
library(R2OpenBUGS)
library(coda)
#Code for OpenBUGS model given below
cat("
model{
    for(i in 1:20)
    {
    #smoke   age   death  pyears 
    
    death[i]~dpois(lam[i])
    log(lam[i]) <- log(pyears[i]) + beta0 + beta.s[smoke[i]] + beta.c[age[i]] + b[i] 
    b[i] ~dnorm(0,tau)
    b.adj[i] <- b[i] - mean(b[]) 
    }
    for(is in 1:4){
    beta.s[is]~dnorm(0,tau.s)
    beta.s.adj[is] <- beta.s[is] -mean(beta.s[]) 
    }
    for(ic in 1:5){
    beta.c[ic]~dnorm(0,tau.c)
    beta.c.adj[ic] <- beta.c[ic] - mean(beta.c[])
    }
    #  Note: total person-years per categories is less than 115,000
    #  ln(115000) ~  = 11.65....
    #  so rate has to be bigger than 1/115,000 and log(rate) > -11.65... 
    #  so log( base rate) should be between about -12 and 12.
    # 1/12/12 is about .00694
    beta0 ~ dnorm(0, .00694)  
    beta0.adj <- beta0 + mean(b[]) + mean(beta.s[])+ mean(beta.c[])
    # for the <extra poisson variation> ... 
    # assume bounded by very big number... say 1000 times... 
    # so log(1000) is about 2.3*4 which is about 9.2
    std ~ dunif(0, 9)
    tau <- 1/std/std
    # for the relative risk between groups... a very large number would be 100 times, 
    #  so, log(100) is about 2.3*2 or about 4.6
    #  also, note that 1/5/5 is 0.04
    #  
    std.s ~dunif(0, 5)
    tau.s <- 1/std.s/std.s
    std.c ~ dunif(0,5) 
    tau.c <- 1/std.c/std.c
    beta.o~dnorm(0, .04)
    }", file="smokemod.txt")

#defining parameters and data for the bugs function

params=c("beta.s.adj",  "std.s", "beta.c.adj", "std.c", "beta0.adj", "std")
attach(smokeDat)

bug.dat=list("smoke","age","death", "pyears")
init.fun=function(){list(
  beta.s=rnorm(4), std.s=runif(1,1,2),
  beta.c=rnorm(5), std.c=runif(1,1,2), 
  std=runif(1,1,2), beta0=rnorm(1),
  b=rnorm(20,0,.1))}

#using openBUGS to run our model, this code also gives the posterior which we need for 2a in OpenBUGS

smokeBug0=bugs(bug.dat, init.fun, params, model.file="smokemod.txt",
              n.chains=5, n.iter=8000, n.burnin=1000, debug=TRUE  #for production
              #	n.chains=5, n.iter=6000, n.burnin=1000, debug=TRUE   #for testing
)

```


openBUGS Results output showing the posterior distributions:


```{r}
library(knitr)
results = read.csv("question2a_results.csv")
kable(results, caption = "Summary Statistics")
#Deviance Information
DevianceInformation = read.csv("DevianceInformation.csv")
kable(DevianceInformation, caption = "Deviance Information")
```


\begin{center} <br>
\includegraphics[width=8in]{question2output.pdf} <br>
\end{center}

\newpage

The above plots show the convergence in the chains after 8000 iterations

```{r}

#boxplots for smoking categories

boxplot(data.frame( (smokeBug0$sims.list)["beta.s.adj"]),
        xlab="beta.s.adj")

#boxplots for age categories

boxplot(data.frame( (smokeBug0$sims.list)["beta.c.adj"]),
        xlab="beta.c.adj")

```


## Question 2 b. 

The following plots show that our model converges.

Trace plots have been shown in the output given above that shows the convergence in chains after 8000 iterations.

The autocorrelation functions plots are given below.


```{r }

# The following plots show that our model converges
#autocorrelation functions given below
#trace plots have been shown in the output given above that shows the convergence in chains after 8000 iterations
acf(smokeBug0$sims.array[,1,"beta.s.adj[1]"]) 
```


After 10 lags the plot seems to have converged. We observe the same for age which converges twice as #fast as shown below

```{r}
#After 10 lags the plot seems to have converged. We observe the same for age which converges twice as #fast as shown below
acf(smokeBug0$sims.array[,1,"beta.c.adj[1]"])

acf(smokeBug0$sims.array[,1,"beta0.adj"])

```


## Question 2 c. 



```{r}
risk20<-smokeBug0$sims.list$beta.s.adj[,4]-smokeBug0$sims.list$beta.s.adj[,1]
mean(risk20)
exp(mean(risk20))

```


The above shows those who smoke >20 cigarettes per day versus nonsmoker have approximately 4 times (=3.98) the risk. We take the exponential in the last line since the model gives us the log of the risks.

```{r}
quantile(exp(risk20), c(.025, .975)) # gives the 95% Cis for above risk
```


# Question 3 


## Question 3 a

R-code given below for defining data parameters, model and priors 

```{r}

library(R2OpenBUGS)

diabetes = read.csv("DiabetesDrugEffect.csv")
attach(diabetes)

## Model 1

cat("
model{
  for( i in 1:12){
   diff[i] ~ dnorm(mu[i], Sediff[i])
   mu[i] ~ dnorm(theta, sd)
   }
  theta ~ dnorm(0, .1) # so, sd =5.  exp(5) ~ 148 which is huge
  sd ~ dnorm(0, .1) 
  }
  ", file="diabetesMod3.txt")
  
bugM3.dat=list( "diff", "Sediff")  # what variable you need in the model

initM3.fun=function(){ list( theta=rnorm(1) ,
                             sd = rnorm(1), 
                             mu = rnorm(1)
                             
    ) }

	
paramsM3=c("mu", "theta", "sd")
     ### what variables you want to monitor
```




```{r, eval=F}

#### Could change the code below...
diabetesBaseM3=bugs(bugM3.dat, initM3.fun, paramsM3, model.file="diabetesMod3.txt",
    n.chains=3, n.iter=3000, n.burnin=500,
    n.thin=1 , debug=TRUE
        )

print(diabetesBaseM3,dig=3)


SArray= diabetesBaseM3$sims.array
vname=attr(SArray,"dimnames")[3][[1]]
chainL=attr(SArray,"dim")[1][[1]]
for(i in 1:length(vname)){ 
    nn=vname[i]
    plot(density(SArray[,,nn]), main=nn)
    xnul=locator(1)    
    acf( SArray[,1,nn], main=nn)  #note: this is only for 1st chain
    xnul=locator(1)
    matplot(1:chainL,SArray[,,nn], main=nn,xlab="index",type="l")
    xnul=locator(1)
}

```



```{r}
##  Model 2

cat("
model{
  for( i in 1:12){
   diff[i] ~ dnorm(theta, Sediff[i] + sd)
   }
  theta ~ dnorm(0, .1) 
  sd ~ dnorm(0, .1) 
  }
  ", file="diabetesMod2.txt")
  
bugM3.dat=list( "diff", "Sediff")  # what variable you need in the model

initM2.fun=function(){ list( theta=rnorm(1) ,
                             sd = rnorm(1)
    ) }

	
paramsM2=c( "theta", "sd")
     ### what variables you want to monitor


```

```{r, eval = F}
#### Could change the code below...
diabetesBaseM2=bugs(bugM3.dat, initM2.fun, paramsM2, model.file="diabetesMod2.txt",
    n.chains=3, n.iter=3000, n.burnin=500,
    n.thin=1 , debug=TRUE
        )

print(diabetesBaseM2,dig=3)

SArray= diabetesBaseM2$sims.array
vname=attr(SArray,"dimnames")[3][[1]]
chainL=attr(SArray,"dim")[1][[1]]
for(i in 1:length(vname)){ 
    nn=vname[i]
    plot(density(SArray[,,nn]), main=nn)
    xnul=locator(1)    
    acf( SArray[,1,nn], main=nn)  #note: this is only for 1st chain
    xnul=locator(1)
    matplot(1:chainL,SArray[,,nn], main=nn,xlab="index",type="l")
    xnul=locator(1)
}
```

